---
title: "[Hands-On Exercise 8-3:]{style='color:DarkOrange'} Analytical Mapping"
date: 2025-06-11
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# [8-3:]{style="color:DarkOrange"} 1 Overview

## [8-3:]{style="color:DarkOrange"} 1.1 Objectives

This exercise will cover using appropriate R methods to plot analytical maps.

## [8-3:]{style="color:DarkOrange"} 1.2 Learning Outcomes

The following functions of tmap and tidyverse will be explored:

-   Importing geospatial data in rds format into R environment
-   Creating cartographic quality choropleth maps by using appropriate tmap functions
-   Creating rate map
-   Creating percentile map
-   Creating boxmap

# [8-3:]{style="color:DarkOrange"} 2 Getting Started

## [8-3:]{style="color:DarkOrange"} 2.1 Installing and Loading Packages

```{r}
#| code-fold: true
#| code-summary: "Show Code"
pacman::p_load(sf, tmap, tidyverse, patchwork)
```

## [8-3:]{style="color:DarkOrange"} 2.1 Importing Data to R

For the purpose of this hands-on exercise, a prepared data set called *NGA_wp.rds* will be used. The data set is a polygon feature data.frame providing information on water point of Nigeria at the LGA level.

The code below importa NGA_wp using *read_rds()*.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

# [8-3:]{style="color:DarkOrange"} 3 Basic Choropleth Mapping

## [8-3:]{style="color:DarkOrange"} 3.1 **Visualising distribution of non-functional water point**

To plot a choropleth map showing the distribution of non-function water point by LGA

::: panel-tabset

## Original
```{r}
#| code-fold: true
#| code-summary: "Show Code"
p1 <- tm_shape(NGA_wp) +
  tm_polygons(fill = "wp_functional",
             fill.scale = tm_scale_intervals(
               style = "equal",
               n = 10,
               values = "brewer.blues"),
             fill.legend = tm_legend(
               position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Distribution of functional water point by LGAs")

p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp", 
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + 
  tm_title("Distribution of total  water point by LGAs")

tmap_arrange(p2, p1, nrow = 1)
```

## Improved
```{r}
#| code-fold: true
#| code-summary: "Show Code"
p1 <- tm_shape(NGA_wp) +
  tm_polygons(fill = "wp_functional",
             fill.scale = tm_scale_intervals(
               style = "equal",
               n = 10,
               values = "brewer.blues"),
             fill.legend = tm_legend(
               position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Distribution of functional water point by LGAs") +
  tm_layout(legend.width = 8,
            legend.height = 8,
            legend.text.size = 0.7)

p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp", 
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + 
  tm_title("Distribution of total  water point by LGAs") +
  tm_layout(legend.width = 8,
            legend.height = 8,
            legend.text.size = 0.7)

tmap_arrange(p2, p1, nrow = 1)
```

:::

# [8-3:]{style="color:DarkOrange"} 4 Choropleth Map for Rates

In many readings, there is importance in mapping rates rather than counting things, and that is for the simple reason that water points are not equally distributed in space. That means that if the number and location of water points are not accounted for, it ends up in mapping the total water point size rather than the topic of interest.

## [8-3:]{style="color:DarkOrange"} 4.4 Deriving Proportion of Functional Water Points and Non-Functional Water Points

The proportion of functional water points and the proportion of non-functional water points in each LGA will be tabulated. In the following code chunk, `mutate(`) from **dplyr** package is used to derive two fields, namely *pct_functional* and *pct_nonfunctional*.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

## [8-3:]{style="color:DarkOrange"} 4.5 Plotting Map of Rate

Plotting a choropleth map showing the distribution of percentage functional water point by LGA.

::: panel-tabset

## Original
```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(NGA_wp) +
  tm_polygons("pct_functional",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) + 
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs")
```

## Improved
```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(NGA_wp) +
  tm_polygons("pct_functional",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) + 
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs") +
  tm_layout(legend.width = 7,
            legend.height = 8,
            legend.text.size = 0.7)
```
:::

# [8-3:]{style="color:DarkOrange"} 5 Extreme Value Maps

Extreme value maps are variations of common choropleth maps where the classification is designed to highlight extreme values at the lower and upper end of the scale, with the goal of identifying outliers.

These maps were developed in the spirit of spatializing EDA, i.e., adding spatial features to commonly used approaches in non-spatial EDA (Anselin 1994).

## [8-3:]{style="color:DarkOrange"} 5.1 Percentile Map

The percentile map is a special type of quantile map with six specific categories: 0-1%, 1-10%, 10-50%, 50-90%, 90-99%, and 99-100%. 

The corresponding breakpoints can be derived by means of the base R quantile command, passing an explicit vector of cumulative probabilities as c(0,.01,.1,.5,.9,.99,1). 

::: callout-note
The beginning and endpoint must be included
:::

### [8-3:]{style="color:DarkOrange"} 5.1.1 Data Preparation

Step 1: Exclude records with NA by using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
NGA_wp <- NGA_wp %>%
  drop_na()
```

Step 2: Create customised classification and extracting values

```{r}
#| code-fold: true
#| code-summary: "Show Code"
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

::: callout-warning
- When variables are extracted from an sf data.frame, the geometry is extracted as well
- For mapping and spatial manipulation, this is the expected behaviour, but many base R functions cannot deal with the geometry
- Specifically, the quantile() gives an error
- As a result st_set_geomtry(NULL) is used to drop geometry field.
:::

### [8-3:]{style="color:DarkOrange"} 5.1.2 Why Write Functions?

Writing a function has three big advantages over using copy-and-paste:

-   A function can have an evocative name that makes the **code easier to understand**
-   As requirements change, the code only needs to be **updated code in one place**, instead of many
-   Eliminates the chance of making incidental mistakes during copy and paste (i.e. updating a variable name in one place, but not in another)

Source: [Chapter 19: Functions](https://r4ds.had.co.nz/functions.html#functions) of **R for Data Science**.

### [8-3:]{style="color:DarkOrange"} 5.1.3 Creating the get.var Function

Firstly, an R function is written as shown below to extract a variable (i.e. *wp_nonfunctional*) as a vector out of an sf data.frame.

-   arguments:
    -   vname: variable name (as character, in quotes)
    -   df: name of sf data frame
-   returns:
    -   v: vector with values (without a column name)

```{r}
#| code-fold: true
#| code-summary: "Show Code"
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

### [8-3:]{style="color:DarkOrange"} 5.1.4 A Percentile Mapping Function

Next, a percentile mapping function is written using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_polygons(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```

### [8-3:]{style="color:DarkOrange"} 5.1.5 Test the Percentile Mapping Function

The code chunk below is used to run the function.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
percentmap("total_wp", NGA_wp)
```

::: callout-note
This is just a bare bones implementation. Additional arguments such as the title, legend positioning just to name a few of them, could be passed to customise various features of the map.
:::

### [8-3:]{style="color:DarkOrange"} 5.1.6 Improvement

```{r}
#| code-fold: true
#| code-summary: "Show Code"
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_polygons(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle,
            main.title.position = "center",
            frame = FALSE)
}
  
percentmap("total_wp", NGA_wp)
```

## [8-3:]{style="color:DarkOrange"} 5.2 Box Map

In essence, a box map is an augmented quartile map, with an additional lower and upper category. When there are lower outliers, then the starting point for the breaks is the minimum value, and the second break is the lower fence. 

In contrast, when there are no lower outliers, then the starting point for the breaks will be the lower fence, and the second break is the minimum value (there will be no observations that fall in the interval between the lower fence and the minimum value).

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 5
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()
```

::: callout-tip
-   Displays summary statistics of a choropleth map by using the basic principles of boxplot.
-   To create a box map, a custom breaks specification must be used. However, there is a complication. The break points for the box map vary depending on whether lower or upper outliers are present.
:::

### [8-3:]{style="color:DarkOrange"} 5.2.1 Creating the boxbreaks Function

The code chunk below is an R function that creates break points for a box map.

-   arguments:
    -   v: vector with observations
    -   mult: multiplier for IQR (default 1.5)
-   returns:
    -   bb: vector with 7 break points compute quartile and fences

```{r}
#| code-fold: true
#| code-summary: "Show Code"
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

### [8-3:]{style="color:DarkOrange"} 5.2.2 Creating the get.var Function

The code chunk below is an R function to extract a variable as a vector out of a sf data frame.

-   arguments:
    -   vname: variable name (as character, in quotes)
    -   df: name of sf data frame
-   returns:
    -   v: vector with values (without a column name)

```{r}
#| code-fold: true
#| code-summary: "Show Code"
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

### [8-3:]{style="color:DarkOrange"} 5.2.3 Test the Newly Created Function

The code chunk below runs the newly created function.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

### [8-3:]{style="color:DarkOrange"} 5.2.4 Boxmap function

The code chunk below is an R function to create a box map.

-   arguments:
    -   vnam: variable name (as character, in quotes)
    -   df: simple features polygon layer
    -   legtitle: legend title
    -   mtitle: map title
    -   mult: multiplier for IQR -
-   returns:
    -   a tmap-element (plots a map)

::: panel-tabset

## Original
```{r}
#| code-fold: true
#| code-summary: "Show Code"
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}

tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```

## Improvement
```{r}
#| code-fold: true
#| code-summary: "Show Code"
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            main.title.position = c("center"),
            frame = FALSE)
}

tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```
:::