---
title: "[Take-Home Exercise 2:]{style='color:DarkOrange'} VAST Mini-Challenge 1 2025 (Sailor Shift and Oceanus Folk)"
date: 2025-05-15
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# [2:]{style="color:DarkOrange"} 1 Overview

This exercise will target to answer one of the three challenges from [VAST 2025](https://vast-challenge.github.io/2025/index.html) which features a fictitious island nation, Oceanus, famous for fishing and quiet seaside communities.

The chosen challenge is [Mini-Challenge 1](https://vast-challenge.github.io/2025/MC1.html).

## [2:]{style="color:DarkOrange"} 1.1 Background

One of music’s biggest superstars is Oceanus native Sailor Shift. From humble beginnings, Sailor has grown in popularity and now enjoys fans around the world. Sailor started her career on the island nation of Oceanus which can be clearly seen in her early work, she started in the genre of “Oceanus Folk”. While Sailor has moved away from the traditional Oceanus style, the Oceanus Folk has made a name for itself in the musical world. The popularity of this music is one of the factors driving an increase in tourism to a quiet island nation that used to be known for fishing.

In 2023, Sailor Shift joined the Ivy Echoes – an all-female Oceanus Folk band consisting of Sailor (vocalist), Maya Jensen (vocalist), Lila “Lilly” Hartman (guitarist), Jade Thompson (drummer), and Sophie Ramirez (bassist). They played together at venues throughout Oceanus but had broken up to pursue their individual careers by 2026. Sailor’s breakthrough came in 2028 when one of her singles went viral, launched to the top of the global charts (something no other Oceanus Folk song had ever done). Since then, she has only continued to grow in popularity worldwide.

Sailor has released a new album almost every year since her big break, and each has done better than the last. Although she has remained primarily a solo artist, she has also frequently collaborated with other established artists, especially in the Indie Pop and Indie Folk genres. She herself has branched out musically over the years but regularly returns to the Oceanus Folk genre — even as the genre’s influence on the rest of the music world has spread even more.

Sailor has always been passionate about two things: (1) spreading Oceanus Folk, and (2) helping lesser-known artists break into music. Because of those goals, she’s particularly famous for her frequent collaborations.

Additionally, because of Sailor’s success, more attention began to be paid over the years to her previous bandmates. All 4 have continued in the music industry—Maya as an independent vocalist, Lilly and Jade as instrumentalists in other bands, and Sophie as a music producer for a major record label. In various ways, all of them have contributed to the increased influence of Oceanus folk, resulting in a new generation of up-and-coming Oceanus Folk artists seeking to make a name for themselves in the music industry.

Now, as Sailor returns to Oceanus in 2040, a local journalist – Silas Reed – is writing a piece titled Oceanus Folk: Then-and-Now that aims to trace the rise of Sailor and the influence of Oceanus Folk on the rest of the music world. He has collected a large dataset of musical artists, producers, albums, songs, and influences and organized it into a knowledge graph. Your task is to help Silas create beautiful and informative visualizations of this data and uncover new and interesting information about Sailor’s past, her rise to stardom, and her influence.

## [2:]{style="color:DarkOrange"} 1.2 Tasks and Questions

1.  Design and develop visualizations and visual analytic tools that will allow Silas to explore and understand the profile of Sailor Shift’s career

    -   Who has she been most influenced by over time?
    -   Who has she collaborated with and directly or indirectly influenced?
    -   How has she influenced collaborators of the broader Oceanus Folk community?

2.  Develop visualizations that illustrate how the influence of Oceanus Folk has spread through the musical world.

    -   Was this influence intermittent or did it have a gradual rise?
    -   What genres and top artists have been most influenced by Oceanus Folk?
    -   On the converse, how has Oceanus Folk changed with the rise of Sailor Shift? From which genres does it draw most of its contemporary inspiration?

3.  Use your visualizations to develop a profile of what it means to be a rising star in the music industry.

    -   Visualize the careers of three artists. Compare and contrast their rise in popularity and influence.
    -   Using this characterization, give three predictions of who the next Oceanus Folk stars with be over the next five years.
    
## [2:]{style="color:DarkOrange"} 1.3 Data Source

The data for this exercise is from [VAST 2025 MC1](https://github.com/vast-challenge/2025-data/blob/main/MC1_release.zip). 

## [2:]{style="color:DarkOrange"} 1.4 Data Description

Graph Description
-	Directed Multigraph
-	# nodes: 17,412
-	# edges: 37, 857
-	18 connected components
-	Possible node types are: {Person, Song, RecordLabel, Album, MusicalGroup}
-	Possible edge types are: {MemberOf, PerformerOf, ComposerOf, ProducerOf, LyricistOf, InStyleOf, InterpolatesFrom, CoverOf, LyricalReferenceTo, DirectlySamples, RecordedBy, DistributedBy}
-	The graph format is a JSON file. The root-level JSON object consists of graph-level properties specifying that it is directed and a multigraph, a “nodes” key which holds the list of nodes, and a “links” key which holds the list of edges.

The data for this challenge comes from two different sources:

-	an online, crowdsourced repository of musical influence, where contributors have  manually notated instances in which songs or albums have sampled, covered, or otherwise drawn inspiration from previous work. 
-	aggregate-level descriptions of song popularity, as labeled by our journalist, Silas Reed. These notations (notable and notoriety_date) are directly based on both the number of sales and streams that the song/album has achieved as well as the whether or not the work landed on a top record chart.

## [1-1:]{style="color:DarkOrange"} 1.5 Install and Load the Required Libraries

Utility Tools

- jsonlite: To parse JSON
- tidyverse - Data science tools
- ggtext - Tools for text formatting
- knitr - For better table displays
- lubridate - For processing date and time

Graphing Tools

- patchwork - For combining ggplot plots
- ggraph - For plotting network data
- tidygraph - For graph manipulations
- igraph - Contains functions for network analysis
- ggiraph - Interactive plots

::: {.callout-note appearance="minimal" collapse="true"}
## Code

```{r}
#| code-fold: true
#| code-summary: "Show code"
pacman::p_load(jsonlite, tidyverse, ggtext,
                knitr, lubridate, patchwork,
                ggraph, tidygraph, igraph,
                ggiraph)
```
:::

## [1-1:]{style="color:DarkOrange"} 1.6 Importing Data

For the purpose of this exercise, a data file called *MC1_graph* will be used. The code chunk below imports *respopagesex2024.csv* into R environment by using *read_csv()* function of *readr* package (under *tidyverse*).

::: {.callout-note appearance="minimal" collapse="true"}
## Code

```{r}
#| code-fold: true
#| code-summary: "Show code"
mc1_data <- fromJSON("data/MC1_graph.json")
glimpse(mc1_data)
```
:::

# [2:]{style="color:DarkOrange"} 2 Exploratory Data Analysis (EDA)

```{r}
mc1_nodes_raw <- as_tibble(mc1_data$nodes)

glimpse(mc1_nodes_raw)

kable(head(mc1_nodes_raw, 5))
```

```{r}
mc1_nodes_raw$release_date %>% unique()

mc1_nodes_raw %>%
  filter(grepl("Sailor Shift", name))
```


```{r}
mc1_nodes_clean <- mc1_nodes_raw %>%
  mutate(
    released_year = as.integer(release_date),
    name = gsub("'", "", name)) %>%
  select(`Node Type`, name, released_year, genre, notable)
  
kable(head(mc1_nodes_clean))
```


```{r}
mc1_edges_raw <- as_tibble(mc1_data$links)

glimpse(mc1_edges_raw)

kable(head(mc1_edges_raw, 10))
```

```{r}
range(mc1_edges_raw$source)

range(mc1_edges_raw$target)
```


```{r}
mc1_edges_format <- merge(mc1_edges_raw, mc1_nodes_raw, by.x = "source", by.y = "id", all.x = TRUE) %>%
  rename(from = name) %>%
  select(from, target, `Edge Type`)
  
mc1_edges_format <- merge(mc1_edges_format, mc1_nodes_raw, by.x = "target", by.y = "id", all.x = TRUE) %>%
  rename(to = name) %>%
  select(from, to, `Edge Type`)

kable(head(mc1_edges_format))
```


```{r}
mc1_edges_clean <- mc1_edges_format %>%
  group_by(`Edge Type`, from, to) %>%
  summarize(weight = n())

glimpse(mc1_edges_clean)

kable(head(mc1_edges_clean))
```

```{r}
range(mc1_edges_clean$weight)
```

```{r}
mc1_edges_clean <- mc1_edges_raw %>%
  rename(from = source, to = target) %>%
  mutate(from = from + 1,
         to = to + 1) %>%
  select(`Edge Type`, from, to)

glimpse(mc1_edges_clean)

kable(head(mc1_edges_clean))
```

```{r}
mc1_nodes_clean %>%
  filter(!is.na(released_year))
```


```{r}
music = tbl_graph(edges = mc1_edges_clean,
                             nodes = mc1_nodes_clean,
                             directed = TRUE)
```

## 1.1

```{r}
# Convert tbl_graph to igraph
ig <- as.igraph(music)

# Get index for 'Sailor Shift'
sailor_idx <- which(V(ig)$name == "Sailor Shift")

# Get 2-degree ego neighbourhood
neighbourhood_nodes <- ego(ig, order = 3, nodes = sailor_idx, mode = "all")[[1]]

# Induce subgraph and convert to tbl_graph
sub_music <- induced_subgraph(ig, vids = neighbourhood_nodes) %>% 
  as_tbl_graph()

# Convert to igraph for path computation
ig_sub <- as.igraph(sub_music)

# Re-identify 'Sailor Shift' in the new graph
sailor_idx_sub <- which(V(ig_sub)$name == "Sailor Shift")

# Compute shortest paths from Sailor Shift
sp <- shortest_paths(ig_sub, from = sailor_idx_sub, mode = "out", output = "vpath")$vpath

# Count path dependencies (with safety for unreachable nodes)
node_names <- V(ig_sub)$name
dep_score <- sapply(node_names, function(n) {
  sum(sapply(sp, function(path) {
    if (length(path) == 0) return(FALSE)
    n %in% names(path)
  }))
})

# Attach to tbl_graph
sub_music <- sub_music %>%
  mutate(sailor_dependency = dep_score)
```

```{r}
g <- sub_music %>%
  ggraph(layout = "nicely") + 
  geom_edge_fan(
    aes(
      colour = `Edge Type`,
      start_cap = circle(1, 'mm'),
      end_cap = circle(1, 'mm')
      ),
    arrow = arrow(length = unit(2, 'mm')),
    alpha = 0.4
    ) +
  geom_point_interactive(
    aes(
      x = x,
      y = y,
      data_id = name,
      tooltip = sprintf("%s<br/>(%s)", name, released_year),
      colour = genre,
      size = sailor_dependency,
      shape = `Node Type`
      ),
    show.legend = c(size = FALSE)
    )+ 
  geom_node_text(
    aes(
      label = ifelse(`name` == "Sailor Shift", "Sailor Shift", NA)
    ),
    size = 4,
    colour = 'red',
    show.legend = FALSE
    ) +
  theme_graph() +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 9)) +
  scale_shape_discrete(name = "Node Type")

girafe(ggobj = g, width_svg = 10, height_svg = 9)
```

```{r}
# Extract node data
node_data <- as_tibble(sub_music)

# View top 5 nodes with highest sailor_dependency
node_data %>%
  arrange(desc(sailor_dependency)) %>%
  head(5) %>%
  select(`Node Type`, name, sailor_dependency) %>%
  kable()
```

## 1.2  Direct

```{r}
# Convert tbl_graph to igraph
ig <- as.igraph(music)

# Get index for 'Sailor Shift'
sailor_idx <- which(V(ig)$name == "Sailor Shift")

# Get 2-degree ego neighbourhood
neighbourhood_nodes <- ego(ig, order = 1, nodes = sailor_idx, mode = "all")[[1]]

# Induce subgraph and convert to tbl_graph
sub_music <- induced_subgraph(ig, vids = neighbourhood_nodes) %>% 
  as_tbl_graph()

# Convert to igraph for path computation
ig_sub <- as.igraph(sub_music)

# Re-identify 'Sailor Shift' in the new graph
sailor_idx_sub <- which(V(ig_sub)$name == "Sailor Shift")

# Compute shortest paths from Sailor Shift
sp <- shortest_paths(ig_sub, from = sailor_idx_sub, mode = "out", output = "vpath")$vpath

# Count path dependencies (with safety for unreachable nodes)
node_names <- V(ig_sub)$name
dep_score <- sapply(node_names, function(n) {
  sum(sapply(sp, function(path) {
    if (length(path) == 0) return(FALSE)
    n %in% names(path)
  }))
})

# Attach to tbl_graph
sub_music <- sub_music %>%
  mutate(sailor_dependency = dep_score)
```

```{r}
g <- sub_music %>%
  ggraph(layout = "nicely") + 
  geom_edge_fan(
    aes(
      colour = `Edge Type`,
      start_cap = circle(2, 'mm'),
      end_cap = circle(3, 'mm')
      ),
    arrow = arrow(length = unit(2, 'mm')),
    alpha = 0.4
    ) +
  geom_point_interactive(
    aes(
      x = x,
      y = y,
      data_id = name,
      tooltip = sprintf("%s<br/>(%s)", name, released_year),
      colour = genre,
      size = sailor_dependency,
      shape = `Node Type`
      ),
    show.legend = c(size = FALSE)
    )+ 
  geom_node_text(
    aes(
      label = ifelse(`name` == "Sailor Shift", "Sailor Shift", NA)
    ),
    size = 4,
    colour = 'red',
    show.legend = FALSE
    ) +
  theme_graph() +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 9)) +
  scale_shape_discrete(name = "Node Type")

girafe(ggobj = g, width_svg = 8, height_svg = 6)
```

```{r}
# Extract node data
node_data <- as_tibble(sub_music)

node_data %>%
  filter(`Node Type` %in% c("Person", "MusicalGroup")) %>%
  arrange(desc(sailor_dependency)) %>%
  select(`Node Type`, name, sailor_dependency) %>%
  kable()
```

## 1.2  inDirect

```{r}
# Convert tbl_graph to igraph
ig <- as.igraph(music)

# Step 1: Get Sailor Shift's 1-degree out-neighbours that are Songs or Albums
sailor_idx <- which(V(ig)$name == "Sailor Shift")
out_neighbours_all <- neighbors(ig, sailor_idx, mode = "all")

# Filter to keep only nodes of type Song or Album
out_neighbours <- out_neighbours_all[
  V(ig)[out_neighbours_all]$`Node Type` %in% c("Song", "Album")
]

# Step 2: Get 2-degree in-neighbours to those filtered out-neighbours
in_to_out_neighbours <- unique(unlist(
  lapply(out_neighbours, function(n) {
    ego(ig, order = 1, nodes = n, mode = "in")[[1]]
  })
))

# Step 3: Combine all relevant nodes (Sailor Shift + out-neighbours + their 2-degree in-neighbours)
all_nodes <- unique(c(sailor_idx, out_neighbours, in_to_out_neighbours))

# Step 4: Induce subgraph over combined nodes
subgraph <- induced_subgraph(ig, vids = all_nodes)
sub_music <- as_tbl_graph(subgraph)

# Convert to igraph for path computation
ig_sub <- as.igraph(sub_music)

# Re-identify 'Sailor Shift' in the new graph
sailor_idx_sub <- which(V(ig_sub)$name == "Sailor Shift")

# Compute shortest paths from Sailor Shift (all directions in subgraph)
sp <- shortest_paths(ig_sub, from = sailor_idx_sub, mode = "all", output = "vpath")$vpath

# Count path dependencies (including Sailor Shift)
node_names <- V(ig_sub)$name
dep_score <- sapply(node_names, function(n) {
  sum(sapply(sp, function(path) {
    if (length(path) == 0) return(FALSE)
    n %in% names(path)
  }))
})

# Attach path dependency score to graph
sub_music <- sub_music %>%
  mutate(sailor_dependency = dep_score)
```

```{r}
g <- sub_music %>%
  ggraph(layout = "nicely") + 
  geom_edge_fan(
    aes(
      colour = `Edge Type`,
      start_cap = circle(2, 'mm'),
      end_cap = circle(3, 'mm')
      ),
    arrow = arrow(length = unit(2, 'mm')),
    alpha = 0.4
    ) +
  geom_point_interactive(
    aes(
      x = x,
      y = y,
      data_id = name,
      tooltip = sprintf("%s<br/>(%s)", name, released_year),
      colour = genre,
      size = sailor_dependency,
      shape = `Node Type`
      ),
    show.legend = c(size = FALSE)
    )+ 
  geom_node_text(
    aes(
      label = ifelse(`name` == "Sailor Shift", "Sailor Shift", NA)
    ),
    size = 4,
    colour = 'red',
    show.legend = FALSE
    ) +
  theme_graph() +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 9)) +
  scale_shape_discrete(name = "Node Type")

girafe(ggobj = g, width_svg = 8, height_svg = 6)
```

```{r}
# Extract node data
node_data <- as_tibble(sub_music)

node_data %>%
  filter(`Node Type` %in% c("Person", "MusicalGroup")) %>%
  arrange(desc(sailor_dependency)) %>%
  select(`Node Type`, name) %>%
  kable()
```

```{r}
# Convert tbl_graph to igraph
ig <- as.igraph(music)

# Get index for 'Sailor Shift'
sailor_idx <- which(V(ig)$name == "Sailor Shift")

# Get 3-degree ego from Sailor Shift (as integer indices)
sailor_ego <- ego(ig, order = 1, nodes = sailor_idx, mode = "out")[[1]]

sailor_ego_idx <- as.integer(sailor_ego)

# Get indices for all Oceanus Folk nodes
oceanus_idx <- which(V(ig)$genre == "Oceanus Folk")

# Get 1-degree inward ego for each Oceanus Folk node
oceanus_ego_list <- ego(ig, order = 1, nodes = oceanus_idx, mode = "in")

# Flatten and convert to integer indices
oceanus_ego_idx <- unique(unlist(lapply(oceanus_ego_list, as.integer)))

# Combine both sets of indices
combined_indices <- unique(c(sailor_ego_idx, oceanus_ego_idx))

# Induce subgraph from combined set
sub_music <- induced_subgraph(ig, vids = combined_indices) %>%
  as_tbl_graph()

# Convert to igraph for path computation
ig_sub <- as.igraph(sub_music)

# Re-identify Sailor Shift index in subgraph
sailor_idx_sub <- which(V(ig_sub)$name == "Sailor Shift")

# Compute shortest paths *to* Sailor Shift (mode = "in")
sp <- shortest_paths(ig_sub, from = sailor_idx_sub, mode = "in", output = "vpath")$vpath

# Count path dependencies
node_names <- V(ig_sub)$name
dep_score <- sapply(node_names, function(n) {
  sum(sapply(sp, function(path) {
    if (length(path) == 0) return(FALSE)
    n %in% names(path)
  }))
})

# Add sailor_dependency to tbl_graph
sub_music <- sub_music %>%
  mutate(sailor_dependency = dep_score)
```

```{r}
g <- sub_music %>%
  ggraph(layout = "nicely") + 
  geom_edge_fan(
    aes(
      colour = `Edge Type`,
      start_cap = circle(1, 'mm'),
      end_cap = circle(1, 'mm')
      ),
    arrow = arrow(length = unit(2, 'mm')),
    alpha = 0.4
    ) +
  geom_point_interactive(
    aes(
      x = x,
      y = y,
      data_id = name,
      tooltip = sprintf("%s<br/>(%s)", name, released_year),
      colour = genre,
      size = sailor_dependency,
      shape = `Node Type`
      ),
    show.legend = c(size = FALSE)
    )+ 
  geom_node_text(
    aes(
      label = ifelse(`name` == "Sailor Shift", "Sailor Shift", NA)
    ),
    size = 4,
    colour = 'red',
    show.legend = FALSE
    ) +
  theme_graph() +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 9)) +
  scale_shape_discrete(name = "Node Type")

girafe(ggobj = g, width_svg = 11, height_svg = 8)
```

```{r}
# Convert to igraph
ig <- as.igraph(music)

# Identify Oceanus Folk node indices
oceanus_idx <- which(V(ig)$genre == "Oceanus Folk")

# Build subgraph using ego network around Oceanus Folk members
neighbourhood_nodes <- ego(ig, order = 1, nodes = oceanus_idx, mode = "in") %>%
  unlist() %>%
  unique()

# Induce subgraph and convert to tbl_graph
sub_music <- induced_subgraph(ig, vids = neighbourhood_nodes) %>%
  as_tbl_graph()

# Convert subgraph to igraph again for shortest path computation
ig_sub <- as.igraph(sub_music)

# Re-identify 'Sailor Shift' in the new graph
sailor_idx_sub <- which(V(ig_sub)$name == "Sailor Shift")

# Compute shortest paths TO Sailor Shift (mode = "in")
sp <- shortest_paths(ig_sub, from = sailor_idx_sub, mode = "in", output = "vpath")$vpath

# Compute dependency: how often each node appears along paths to Sailor Shift
node_names <- V(ig_sub)$name
dep_score <- sapply(node_names, function(n) {
  sum(sapply(sp, function(path) {
    if (length(path) == 0) return(FALSE)
    n %in% names(path)
  }))
})

# Attach to tbl_graph
sub_music <- sub_music %>%
  mutate(sailor_dependency = dep_score)
```

```{r}
g <- sub_music %>%
  ggraph(layout = "nicely") + 
  geom_edge_fan(
    aes(
      colour = `Edge Type`,
      start_cap = circle(1, 'mm'),
      end_cap = circle(1, 'mm')
      ),
    arrow = arrow(length = unit(2, 'mm')),
    alpha = 0.4
    ) +
  geom_point_interactive(
    aes(
      x = x,
      y = y,
      data_id = name,
      tooltip = sprintf("%s<br/>(%s)", name, released_year),
      colour = genre,
      size = sailor_dependency,
      shape = `Node Type`
      ),
    show.legend = c(size = FALSE)
    )+ 
  geom_node_text(
    aes(
      label = ifelse(`name` == "Sailor Shift", "Sailor Shift", NA)
    ),
    size = 4,
    colour = 'red',
    show.legend = FALSE
    ) +
  theme_graph() +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 9)) +
  scale_shape_discrete(name = "Node Type")

girafe(ggobj = g, width_svg = 11, height_svg = 8)
```

































```{r}
# Convert to igraph
ig <- as.igraph(music)

# Identify Oceanus Folk node indices
oceanus_idx <- which(V(ig)$genre == "Oceanus Folk")

# Build subgraph using ego network around Oceanus Folk members
neighbourhood_nodes <- ego(ig, order = 2, nodes = oceanus_idx, mode = "in") %>%
  unlist() %>%
  unique()

# Induce subgraph and convert to tbl_graph
sub_music <- induced_subgraph(ig, vids = neighbourhood_nodes) %>%
  as_tbl_graph()

# Convert subgraph to igraph again for shortest path computation
ig_sub <- as.igraph(sub_music)

# Re-identify Oceanus Folk node indices in subgraph
oceanus_idx_sub <- which(V(ig_sub)$genre == "Oceanus Folk")

# Compute shortest paths TO Oceanus Folk members (mode = "in")
sp <- shortest_paths(ig_sub, from = oceanus_idx_sub, mode = "in", output = "vpath")$vpath

# Compute dependency: how often each node appears along paths to Oceanus Folk
node_names <- V(ig_sub)$name
dep_score <- sapply(node_names, function(n) {
  sum(sapply(sp, function(path) {
    if (length(path) == 0) return(FALSE)
    n %in% names(path)
  }))
})

# Attach to tbl_graph
sub_music <- sub_music %>%
  mutate(oceanus_dependency = dep_score)
```

```{r}
g <- sub_music %>%
  ggraph(layout = "nicely") + 
  geom_edge_fan(
    aes(
      colour = `Edge Type`,
      start_cap = circle(1, 'mm'),
      end_cap = circle(1, 'mm')
      ),
    arrow = arrow(length = unit(2, 'mm')),
    alpha = 0.4
    ) +
  geom_point_interactive(
    aes(
      x = x,
      y = y,
      data_id = name,
      tooltip = sprintf("%s<br/>(%s)", name, released_year),
      colour = genre,
      shape = `Node Type`
      ),
    show.legend = c(size = FALSE)
    ) +
  theme_graph() +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(size = 9)) +
  scale_shape_discrete(name = "Node Type")

girafe(ggobj = g, width_svg = 11, height_svg = 8)
```