---
title: "Hands-on Exercise 1"
author: "Calvin Tan"
---

# Getting Started

## Install and Launch R Packages

The code chunk below uses p_load() of pacman package to check if tidyverse these packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(patchwork, tidyverse)
```

## Importing the Data

For the purpose of this exercise, a data file called Exam_data will be used. It consists of year end examination grades of a cohort of primary 3 students from a local school. It is in csv file format.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Ploting Simple Graphs
## A blank graph

```{r}
ggplot(data=exam_data)
```

## Purpose of aes() function
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```
All aesthetics of a plot are specified in the aes() function call and the code above adds MATHS to the x-axis.

## Plotting a Bar Chart

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

### Use #\| eval: false to show code without plotting graph
```{r}
#| eval: false
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

### Use #\| echo: false to plot graph without showing code
```{r}
#| echo: false
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

## Plotting a Dot Graph

### With y-axis vs without y-axis
```{r}
d1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)

d2 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)

d1 + d2
```
The y scale is not very useful, in fact it is very misleading.

## Ploting a Histogram

### Default vs Modifying Bins and Colour
```{r}
h1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()  

h2 <- ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue") 

h1 + h2
```
Fixing bin size fixes the gap and adding colour makes the plot more visually appealing.

### Adding Gender Sub-group
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```
Using sub-groups allows for bivariate analysis.

## Plotting a Kernal Density Estimate

### Default vs Gender Sub-group
```{r}
k1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density() 

k2 <- ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()

k1 + k2
```


## Plotting a Box Plot

### Default vs Notched
```{r}
b1 <- ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()

b2 <- ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)

b1 + b2
```
Notches are used in box plots to help visually assess whether the medians of distributions differ. If the notches do not overlap, this is evidence that the medians are different.

## Plotting a Violin Plot
```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

## Plotting a Scatterplot
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()  
```

# Plotting More Complex Plots

## Plotting a Combined Boxplot and Scatterplot
```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5) 
```

## Adding Statistic Functions to Plots

### Default
```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot()
```
The boxplots below are incomplete because the positions of the means were not shown.

### Using the stat_summary() method vs Using the geom() method
```{r}
s1 <- ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="red",        
               size=4)

s2 <- ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun="mean",           
             colour="red",          
             size=4)

s1 + s2
```
Both methods achieves the same result by adding a red marker to represent the mean.

## Adding a best fit to a scatterplot

### Default
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point()
```

### Best Fit Curve vs Best Fit Line
```{r}
f1 <- ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(linewidth=0.5)

f2 <- ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5)

f1 + f2
```

## Generating Trellis Plots

### Using facet_wrap()
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```
facet_wrap wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than facet_grid because most displays are roughly rectangular.

### Using facet_grid()
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```
facet_grid() forms a matrix of panels defined by row and column facetting variables. It is most useful when you have two discrete variables, and all combinations of the variables exist in the data.

# Using Coordinates

## Vertical vs Horizontal Bar Chart
```{r}
b3 <- ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()

b4 <- ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()

b3 + b4
```

## Changing the y- and x-axis range

### Default vs Equal y- and x-axis range
```{r}
r1 <- ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5)

r2 <- ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

r1 + r2
```
The default scatterplot on the left is slightly misleading because the y-aixs and x-axis range are not equal. The scatterplot on the right fixes both the y-axis and x-axis range to be from 0-100.

# Using Themes
Themes can make the plots more visually appealing.

## Default vs Grey Theme
```{r}
t1 <- ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()

t2 <- ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()

t1 + t2
```

## Other Themes
```{r}
t3 <- ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()

t4 <- ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()

t3 + t4
```