---
title: "[Hands-On Exercise 10:]{style='color:DarkOrange'} Information Dashboard Design: R methods"
date: 2025-06-24
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# [10:]{style="color:DarkOrange"} 1 Overview

This exercise will cover the following:

-   Create bullet chart by using ggplot2
-   Create sparklines by using ggplot2
-   Build industry standard dashboard by using R Shiny

# [10:]{style="color:DarkOrange"} 2 Getting Started

The following R packages will be used in this exercise:

-   [tidyverse](https://www.tidyverse.org/) provides a collection of functions for performing data science task such as importing, tidying, wrangling data and visualising data. It is not a single package but a collection of modern R packages including but not limited to [readr](https://readr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/), [dplyr](https://dplyr.tidyverse.org/), [ggplot](https://ggplot2.tidyverse.org/), [tibble](https://tibble.tidyverse.org/), [stringr](https://stringr.tidyverse.org/), [forcats](https://forcats.tidyverse.org/) and [purrr](https://purrr.tidyverse.org/).
-   [lubridate](https://r4va.netlify.app/chap31) provides functions to work with dates and times more efficiently.
-   [ggthemes](https://r4va.netlify.app/chap31) is an extension of ggplot2. It provides additional [themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) beyond the basic themes of ggplot2.
-   [gtExtras](https://jthomasmock.github.io/gtExtras/index.html) provides some additional helper functions to assist in creating beautiful tables with [gt](https://gt.rstudio.com/), an R package specially designed for anyone to make wonderful-looking tables using the R programming language.
-   [reactable](https://glin.github.io/reactable/index.html) provides functions to create interactive data tables for R, based on the [React Table](https://github.com/tanstack/table/tree/v7) library and made with [reactR](https://github.com/react-R/reactR).
-   [reactablefmtr](https://kcuilla.github.io/reactablefmtr/index.html) provides various features to streamline and enhance the styling of interactive reactable tables with easy-to-use and highly-customizable functions and themes.

The code chunks below are used to installs and loads the above packages into R.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
pacman::p_load(lubridate, ggthemes, reactable,
               reactablefmtr, gt, gtExtras, tidyverse, patchwork)
```

# [10:]{style="color:DarkOrange"} 3 Importing Microsoft Access database

## [10:]{style="color:DarkOrange"} 3.1 The Data

For the purpose of this study, a personal database in Microsoft Access mdb format called *Coffee Chain* will be used.

## [10:]{style="color:DarkOrange"} 3.2 Importing Database into R

In the code chunk below, `odbcConnectAccess()` of [RODBC](https://cran.r-project.org/web/packages/RODBC/index.html) package is used used to import a database query table into R.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
library(RODBC)
con <- odbcConnectAccess2007('data/Coffee Chain.mdb')
coffeechain <- sqlFetch(con, 'CoffeeChain Query')
write_rds(coffeechain, "data/CoffeeChain.rds")
odbcClose(con)
```

::: callout-note
-   Before running the code chunk, the R system must be changed to the 32bit version
-   This is because the `odbcConnectAccess()` is based on 32bit and not 64bit
:::

## [10:]{style="color:DarkOrange"} 3.3 Data Preparation

The code chunk below is used to import *CoffeeChain.rds* into R.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
coffeechain <- read_rds("data/CoffeeChain.rds")
```

::: callout-note
This step is optional if *coffeechain* is already available in R
:::

Next, the code chunk below is used to aggregate Sales and Budgeted Sales at the Product level.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

## [10:]{style="color:DarkOrange"} 3.4 Bullet Chart in ggplot2

The code chunk below is used to plot the bullet charts using ggplot2 functions.

::: panel-tabset
## Improved

```{r}
#| code-fold: true
#| code-summary: "Show Code"
library(ggplot2)
library(dplyr)

ggplot(product, aes(Product, current)) + 
  # Background targets
  geom_col(aes(y = max(target) * 1.01), fill = "grey85", width = 0.85, show.legend = FALSE) +
  geom_col(aes(y = target * 0.75, fill = "Target (75%)"), width = 0.85) +
  geom_col(aes(y = target * 0.5, fill = "Target (50%)"), width = 0.85) +
  # Actual sales
  geom_col(aes(y = current, fill = "Current"), width = 0.35) +
  # Target mark (red line)
  geom_errorbar(aes(y = target, ymin = target, ymax = target, colour = "Target Mark"),
                width = 0.4, size = 1) +
  # Coordinate flip
  coord_flip() +
  # Labels and theme
  labs(y = "Current Sales", fill = "Bars", colour = "Markers") +
  # Manual legend
  scale_fill_manual(
    values = c("Current" = "black",
               "Target (75%)" = "grey60",
               "Target (50%)" = "grey50",
               "Target (100%)" = "grey85")
  ) +
  scale_colour_manual(
    values = c("Target Mark" = "red")
  ) +
  theme(legend.position = "bottom") +
  scale_y_continuous(labels = scales::comma)
```

## Original

```{r}
#| code-fold: true
#| code-summary: "Show Code"
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```
:::

# [10:]{style="color:DarkOrange"} 4 Plotting sparklines using ggplot2

This section will explain how to plot sparklines by using ggplot2.

## [10:]{style="color:DarkOrange"} 4.1 Preparing the Data

The data is filtered to include only records from 1 Jan 2013 onwards and grouped by Month and Product.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

The code chunk below is used to compute the minimum, maximum and end of the month sales.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

Lastly, the code chunk below is used to compute the 25 and 75 quantiles.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)
```

## [10:]{style="color:DarkOrange"} 4.2 sparklines in ggplot2

The code chunk below plots sparklines using ggplot2.

::: panel-tabset
## Improved

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-height: 15
#| fig-width: 15
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 10))
```

## Original

```{r}
#| code-fold: true
#| code-summary: "Show Code"
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```
:::

# [10:]{style="color:DarkOrange"} 5 Static Information Dashboard Design: gt and gtExtras Methods

In this section covers how to create static information dashboard by using [gt](https://gt.rstudio.com/) and [gtExtras](https://jthomasmock.github.io/gtExtras/index.html) packages. Please refer to the links provided to get an overview of the purposes and functions provided by these two packages.

## [10:]{style="color:DarkOrange"} 5.1 Plotting a Simple Bullet Chart

This section covers how to prepare a bullet chart report by using functions of gt and gtExtras packages.

::: panel-tabset
## Improved

```{r}
#| code-fold: true
#| code-summary: "Show Code"
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538() %>%
  cols_label(current = "Current Sales") 
```

## Original

```{r}
#| code-fold: true
#| code-summary: "Show Code"
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```
:::

# [10:]{style="color:DarkOrange"} 6 Plotting a Simple Bullet Chart

The code chunk below will be used to prepare the data for creating the sales report by product.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()
```

It is important to note that one of the requirement of gtExtras functions is that they require data.framea to be passed with list columns. In view of this, code chunk below will be used to convert the report data.frame into list columns.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

## [10:]{style="color:DarkOrange"} 6.1 Plotting Coffechain Sales report

The code chunk below plots a simple sales report.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)
```

## [10:]{style="color:DarkOrange"} 6.2 Adding Statistics

First, the code chunk below calculates the summary statistics of Sales by Product such as Min, Max and Average.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

## [10:]{style="color:DarkOrange"} 6.3 Combining the data.frame

The code chunk below combines the sales report with the summary statistics report.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")

sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )

sales_data = left_join(sales, spark)

sales_data
```

## [10:]{style="color:DarkOrange"} 6.4 Plotting the Updated data.table

The data.table created above is plotted using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```

## [10:]{style="color:DarkOrange"} 6.5 Combining Bullet Chart and sparklines

Similarly, the bullet chart and sparklines can be combined using the steps below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Current Sales` = sum(`Sales`)) %>%
  ungroup()

sales_data = sales_data %>%
  left_join(bullet)

sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = `Current Sales`, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

# [10:]{style="color:DarkOrange"} 7 Interactive Information Dashboard Design: reactable and reactablefmtr methods

This section will cover how to create interactive information dashboard by using [reactable](https://glin.github.io/reactable/index.html) and [reactablefmtr](https://kcuilla.github.io/reactablefmtr/index.html) packages. Please refer to the links provided to get an overview of the purposes and functions provided by these two packages.

The code chunk below installs dataui R package which is needed to build an interactive sparklines.

::: {.callout-note appearance="minimal" collapse="true"}
## Show Code

``` r
remotes::install_github("timelyportfolio/dataui")
```
:::

Next, the package is loaded onto R environment by using the code chunk below.

::: {.callout-note appearance="minimal" collapse="true"}
## Show Code

```{r}
library(dataui)
```
:::

## [10:]{style="color:DarkOrange"} 7.1 Plotting Interactive sparklines

Similar to gtExtras, the list field has to be prepared to plot an interactive sparklines by using the reactablefmtr package.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

Next, [react_sparkline](https://kcuilla.github.io/reactablefmtr/reference/react_sparkline.html) will be to plot the sparklines as shown below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

## [10:]{style="color:DarkOrange"} 7.2 Changing the pagesize

By default the pagesize is 10. In the code chunk below, arguments defaultPageSize is used to change the default setting.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

## [10:]{style="color:DarkOrange"} 7.3 Adding points and labels

In the code chunk below `highlight_points` argument is used to show the minimum and maximum values points and `label` argument is used to label the first and last values.

::: panel-tabset
## Improved

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last"),
        margin = list(top = 10, right = 30, bottom = 10, left = 30)
        )
    )
  )
)
```

## Original

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```
:::

## [10:]{style="color:DarkOrange"} 7.4 Adding reference line

In the code chunk below `statline` argument is used to show the mean line.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```

## [10:]{style="color:DarkOrange"} 7.5 Adding bandline

Instead adding reference line, bandline can be added by using the `bandline` argument.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```

## [10:]{style="color:DarkOrange"} 7.6 Changing from sparkline to sparkbar

Instead of displaying the values as sparklines, it can be display as sparkbars as shown below.

::: panel-tabset
## Improved

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean",
        margin = list(top = 0, right = 40, bottom = 0, left = 30)
        )
    )
  )
)
```

## Original

```{r}
#| code-fold: true
#| code-summary: "Show Code"
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```
:::
