---
title: "[Hands-On Exercise 9-1:]{style='color:DarkOrange'} Creating Ternary Plot with R"
date: 2025-06-18
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# [9-1:]{style="color:DarkOrange"} 1 Overview

Ternary plots are a way of displaying the distribution and variability of three-part compositional data. For example, the proportion of aged, economy active and young population or sand, silt, and clay in soil.

It displays is a triangle with sides scaled from 0 to 1. Each side represents one of the three components. A point is plotted so that a line drawn perpendicular from the point to each leg of the triangle intersect at the component values of the point.

In this hands-on, s ternary plot will be built programmatically using R for visualising and analysing population structure of Singapore.

It consists of the following four steps:

-   Install and launch **tidyverse** and **ggtern** packages.
-   Derive three new measures using *mutate()* function of **dplyr** package.
-   Build a static ternary plot using *ggtern()* function of **ggtern** package.
-   Build an interactive ternary plot using *plot-ly()* function of **Plotly R** package.

# [9-1:]{style="color:DarkOrange"} 2 Install and Launch R Packages

For this exercise, the following two main R packages will be used:

-   [ggtern](https://www.thebestintech.com/), a ggplot extension specially designed to plot ternary diagrams. The package will be used to plot static ternary plots.
-   [Plotly R](https://plotly.com/r/), an R package for creating interactive web-based graphs via plotly’s JavaScript graphing library, plotly.js . The **plotly R** libary contains the ggplotly function, which will convert **ggplot2** figures into a Plotly object.

The following **tidyvers**e family packages **readr**, **dplyr** and **tidyr** are also installed and loaded.

The code chunk below will accomplish the tasks above.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
pacman::p_load(plotly, ggtern, tidyverse)
```

# [9-1:]{style="color:DarkOrange"} 3 Data Preparation

## [9-1:]{style="color:DarkOrange"} 3.1 Data

The [Singapore Residents by Planning AreaSubzone, Age Group, Sex and Type of Dwelling, June 2000-2018](https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data) data will be used. It is called *respopagsex2000to2018_tidy.csv* and is in the csv file format.

## [9-1:]{style="color:DarkOrange"} 3.2 Import Data

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#Reading the data into R environment
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv")
```

## [9-1:]{style="color:DarkOrange"} 3.3 Prepare Data

The ***mutate()*** function of **dplyr** package is used to derive three new measures, namely: young, active and old.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year)) %>%
  spread(AG, Population) %>% #turn the values in Population col into AG cols.
  mutate(YOUNG = rowSums(.[4:8])) %>% #Age 0-24
  mutate(ACTIVE = rowSums(.[9:16])) %>% #Age 25-64
  mutate(OLD = rowSums(.[17:21])) %>% #Age >65
  mutate(TOTAL = rowSums(.[22:24])) %>% #Age 
  filter(Year == 2018) %>%
  filter(TOTAL >0)
```

# [9-1:]{style="color:DarkOrange"} 4 Plot Ternary Diagram with R

## [9-1:]{style="color:DarkOrange"} 4.1 Plot Static Ternary Diagram

The ***ggtern()*** function of **ggtern** package is used to create a simple ternary plot.

::: panel-tabset
## Default

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#Building the static ternary plot
ggtern(data=agpop_mutated, 
       aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```

## Adding Colour

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#Building the static ternary plot
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```
:::

## [9-1:]{style="color:DarkOrange"} 4.2 Plot Interactive Ternary Diagram

The code below creates an interactive ternary plot using the ***plot_ly()*** function of **Plotly R**.

::: panel-tabset
## Default

```{r}
#| code-fold: true
#| code-summary: "Show Code"
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```

## Improved

```{r}
#| code-fold: true
#| code-summary: "Show Code"
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes,
    title = list(
      text = "<b>Population structure, 2015<b>",
      x = 0.2,  
      xanchor = "center",  # Anchor point for positioning
      y = 0.95
    ),
    margin = list(t = 80)
  )
```
:::
