---
title: "[Hands-On Exercise 8-1:]{style='color:DarkOrange'} Choropleth Mapping with R"
date: 2025-06-11
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# [8-1:]{style="color:DarkOrange"} 1 Overview

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors. For example, a social scientist may need to use a choropleth map to portray the spatial distribution of aged population of Singapore by Master Plan 2014 Subzone Boundary.

This exercise will cover how to plot functional and truthful choropleth maps by using an R package called [**tmap**](https://cran.r-project.org/web/packages/tmap/) package.

::: callout-tip
It is advisable to read the functional description of each function before using them.
:::

# [8-1:]{style="color:DarkOrange"} 2 Getting Started

In this hands-on exercise, the key R package used is [tmap](https://cran.r-project.org/web/packages/tmap/) package in R. Besides the **tmap** package, four other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text files
-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data
-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data
-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data

Among these four packages, **readr, tidyr** and **dplyr** are part of the **tidyverse** package.

The code chunk below will be used to install and load these packages in RStudio.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
pacman::p_load(sf, tmap, tidyverse, patchwork) # patchwork is added to prepare composite figures
```

::: callout-note
Only **tidyverse** needs to be installed instead of the individual **readr, tidyr** and **dplyr** packages.
:::

# [8-1:]{style="color:DarkOrange"} 3 Importing Data into R

## [8-1:]{style="color:DarkOrange"} 3.1 Data

Two data set will be used to create the choropleth map. They are:

-   Master Plan 2014 Subzone Boundary (Web) (i.e. `MP14_SUBZONE_WEB_PL`) in ESRI shapefile format. It can be downloaded at [data.gov.sg](https://data.gov.sg/) This is a geospatial data. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.
-   Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. `respopagesextod2011to2020.csv`). This is an aspatial data fie. It can be downloaded at [Department of Statistics, Singapore](https://www.singstat.gov.sg/). Although it does not contain any coordinates values, its PA and SZ fields can be used as unique identifiers to geocode the `MP14_SUBZONE_WEB_PL` shapefile.

## [8-1:]{style="color:DarkOrange"} 3.2 Importing Geospatial Data into R

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
mpsz <- st_read(dsn = "data/geospatial",
                layer = "MP14_SUBZONE_WEB_PL")
```

The code below examines the data:

```{r}
#| code-fold: true
#| code-summary: "Show Code"
mpsz
```

::: callout-note
-   only the first ten records will be displayed due to the default behaviour of R
-   Use print(mpsz, n = 20) to see more rows.
:::

## [8-1:]{style="color:DarkOrange"} 3.3 Importing attribute data to R

Next, the *respopagsex2011to2020.csv* file will be imported into RStudio and saved as an R dataframe called *popagsex*.

The task will be performed using the `*read_csv()*` function of **readr** package as shown in the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

## [8-1:]{style="color:DarkOrange"} 3.4 Data Preparation

Before a thematic map can be prepared, a data table with year 2020 values is required first. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: Age group 0 to 4 until age group 20 to 24
-   ECONOMY ACTIVE: Age group 25-29 until age group 60-64
-   AGED: Age group 65 and above
-   TOTAL: All age groups
-   DEPENDENCY: Ratio between young and aged against economy active group

## [8-1:]{style="color:DarkOrange"} 3.4.1 Data Wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package
-   *mutate(), filter(), group_by()* and *select()* of **dplyr** package

```{r}
#| code-fold: true
#| code-summary: "Show Code"
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG,
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`,
       `ECONOMY ACTIVE`, `AGED`,
       `TOTAL`, `DEPENDENCY`)
```

## [8-1:]{style="color:DarkOrange"} 3.4.2 Joining the Attribute Data and Geospatial Data

Before performing the georelational join, one extra step is required to convert the values in PA and SZ fields to uppercase. This is because the values of PA and SZ fields are made up of both upper- and lowercase. On the other, hand the *SUBZONE_N* and *PLN_AREA_N* are only in uppercase.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ),
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, left_join() of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ* as the common identifier.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

::: callout-tip
## Thing to learn from the code

*left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.
:::

```{r}
#| code-fold: true
#| code-summary: "Show Code"
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

# [8-1:]{style="color:DarkOrange"} 4 Choropleth Mapping Geospatial Data Using *tmap*

Two approaches can be used to prepare thematic map using tmap, they are:

-   Plotting a thematic map quickly by using *qtm()*.
-   Plottinga highly customisable thematic map by using tmap elements.

## [8-1:]{style="color:DarkOrange"} 4.1 Plotting a Choropleth Map Quickly using *qtm()*

The easiest and quickest to draw a choropleth map using **tmap** is by using *qtm()*. It is concise and provides a good default visualisation in many cases.

The code chunk below draws a cartographic standard choropleth map as shown below.

::: panel-tabset
## Static

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tmap_mode("plot")
qtm(mpsz_pop2020,
    fill = "DEPENDENCY")
```

## Interactive

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tmap_mode("view")
qtm(mpsz_pop2020,
    fill = "DEPENDENCY")
```
:::

::: callout-tip
## Thing to learn from the code

-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, the “view” option should be used instead.
-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)
:::

## [8-1:]{style="color:DarkOrange"} 4.2 Creating a Choropleth Map using *tmap’s* Elements

Despite *qtm()* usefulness of drawing a choropleth map quickly and easily, the disadvantge is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tmap_mode("plot")

tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45,
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\\n and Population data from Department of Statistics DOS",
             position = c("left", "bottom"))
```

The following sub-sections below will cover the tmap functions used to plot the elements above.

### [8-1:]{style="color:DarkOrange"} 4.2.1 Drawing a Base Map

The basic building block of **tmap** is *tm_shape()* followed by one or more layer elements such as *tm_fill()* and *tm_polygons()*.

In the code chunk below, *tm_shape()* is used to define the input data (i.e *mpsz_pop2020*) and *tm_polygons()* is used to draw the planning subzone polygons.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

### [8-1:]{style="color:DarkOrange"} 4.2.2 Drawing a choropleth map using *tm_polygons()*

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, the target variable such as *Dependency* has to be assigned to *tm_polygons()*.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

::: callout-tip
## Things to learn from *tm_polygons()*

-   The default interval binning used to draw the choropleth map is called “pretty”. A detailed discussion of the data classification methods supported by **tmap** will be provided in [Section 4.3]((https://isss608-ay2024-25t3.netlify.app/hands-on_ex/hands-on_ex08-1/hands-on_ex08-1#data-classification-methods-of-tmap).
-   The default colour scheme used is `YlOrRd` of ColorBrewer. More information about this colour scheme can be found in [Section 4.4]((https://isss608-ay2024-25t3.netlify.app/hands-on_ex/hands-on_ex08-1/hands-on_ex08-1#color-scheme).
-   By default, *Missing* value will be shaded in grey.
:::

### [8-1:]{style="color:DarkOrange"} 4.2.3 Drawing a Choropleth Map using *tm_fill()* and *tm_border()*

*tm_polygons()* is a wraper of *tm_fill()* and *tm_border()*. *tm_fill()* shades the polygons by using the default colour scheme and *tm_borders()* adds the borders of the shapefile onto the choropleth map.

The code chunk below draws a choropleth map by using *tm_fill()* alone.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

::: callout-note
The planning subzones are shared according to the respective dependency values
:::

To add the boundary of the planning subzones, *tm_borders* is used as shown in the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  fill_alpha = 1)
```

::: callout-tip
## Things to learn from the code above

-   Light-gray border lines have been added to the choropleth map.
-   The alpha argument is used to define the transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).
-   Beside alpha argument, there are three other arguments for *tm_borders()*, they are:
    -   *col*: Border colour
    -   *lwd*: Border line width. The default is 1
    -   *lty*: Border line type. The default is “solid”
:::

## [8-1:]{style="color:DarkOrange"} 4.3 Data Classification Methods of tmap

Most choropleth maps employ some methods of data classification. The point of classification is to take a large number of observations and group them into data ranges or classes.

**tmap** provides a total ten data classification methods, namely: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.

To define a data classification method, the *style* argument of *tm_fill()* or *tm_polygons()* must be used.

### [8-1:]{style="color:DarkOrange"} 4.3.1 Plotting Choropleth Maps with Built-In Classification Methods

The code chunk below shows a quantile data classification that used 5 classes and an equal data classification method.

::: panel-tabset
## Quantile

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## Equal

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)

```
:::

::: callout-note
The distribution of quantile data classification method is **more evenly distributed** than the equal data classification method
:::

::: callout-warning
Warning: Maps Lie!
:::

#### [8-1:]{style="color:DarkOrange"} 4.3.1.1 DIY 1

Using what you had learned, prepare choropleth maps by using different classification methods supported by tmap and compare their differences.

Styles:

-   fixed: Manually define breakpoints
-   sd: Classify data based on standard deviations from the mean
-   Pretty: Creates "round" breaks that are visually appealing
-   kmeans: Uses K-means clustering to find natural groupings
-   hclust: Uses hierarchical clustering to group similar values
-   bclust: Uses bootstrap-aggregated clustering for stability
-   fisher: Optimizes breaks to minimize within-class variance (similar to Jenks)

::: panel-tabset
## Fixed

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "fixed",
          breaks = c(0,5,10,15,20)) +
  tm_borders(alpha = 0.5)
```

## Standard Deviation

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "sd") +
  tm_borders(alpha = 0.5)
```

## Pretty

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "pretty") +
  tm_borders(alpha = 0.5)
```

## kmeans

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "kmeans") +
  tm_borders(alpha = 0.5)
```

## hclust

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "hclust") +
  tm_borders(alpha = 0.5)
```

## bclust

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "bclust") +
  tm_borders(alpha = 0.5)
```

## fisher

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "fisher") +
  tm_borders(alpha = 0.5)
```
:::

::: callout-tip
Different styles give different classification results and it depends on the purpose of the graph.
:::

#### [8-1:]{style="color:DarkOrange"} 4.3.1.2 DIY 2

Preparing choropleth maps by using similar classification method but with different numbers of classes (i.e. 2, 6, 10, 20). Compare the output maps, what observation can you draw?

::: panel-tabset
## n=2

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 2,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## n=6

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## n=10

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 10,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

## n=20

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 20,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```
:::

::: callout-tip
Increasing the number of classes results in more granular colour differences till about n = 10. After that, further increasing it to 20 does not have a significant impact.
:::

### [8-1:]{style="color:DarkOrange"} 4.3.2 Plotting Choropleth Map with Custom Breaks

For all the built-in styles, the category breaks are computed internally. In order to override these defaults, the breakpoints can be set explicitly by means of the *breaks* argument to the *tm_fill()*. It is important to note that, in **tmap** the breaks include a minimum and maximum. As a result, in order to end up with n categories, n+1 elements must be specified in the *breaks* option (the values must be in increasing order).

Before starting, it is always a good practice to get some descriptive statistics on the target variable before setting the break points. The code chunk below will be used to compute and display the descriptive statistics of the **DEPENDENCY** field.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
summary(mpsz_pop2020$DEPENDENCY)
```

With reference to the results above, the break points can be set at 0.60, 0.70, 0.80, and 0.90. In addition, the minimum can be set at 0 and the maximum at 100. Therefore, breaks vector will be c(0, 0.60, 0.70, 0.80, 0.90, 1.00).

Now, the choropleth map will be plotted using the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

## [8-1:]{style="color:DarkOrange"} 4.4 Color Scheme

**tmap** supports colour ramps either defined by the user or a set of predefined colour ramps from the **RColorBrewer** package.

### [8-1:]{style="color:DarkOrange"} 4.4.1 Using ColourBrewer Palette

To change the colour, the preferred colour can be assigned to the *values* argument of *tm_scale_intervals()* as shown in the code chunk below.

To reverse the colour shading, a “-” prefix can be added.

::: panel-tabset
## Normal

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

## Reversed

```{r}
#| code-fold: true
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```
:::

::: callout-note
-   Notice that the choropleth map is shaded in green
-   The colour scheme has been reversed in the second tab.
:::

## [8-1:]{style="color:DarkOrange"} 4.5 Map Layouts

Map layout refers to the combination of all map elements into a cohesive map. Map elements include among others the objects to be mapped, the title, the scale bar, the compass, margins and aspects ratios. Colour settings and data classification methods covered in the previous section relate to the palette and break-points and affect how the map looks.

### [8-1:]{style="color:DarkOrange"} 4.5.1 Map Legend

In **tmap**, several *tm_legend()* options are provided to change the placement, format and appearance of the legend.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5,
        values = "brewer.greens"),
      fill.legend = tm_legend(
        title = "Dependency ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of Dependency Ratio by planning subzone \n(Jenks classification)")
```

### [8-1:]{style="color:DarkOrange"} 4.5.2 Map Style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.

The code chunk below shows the *classic* style is used.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

### [8-1:]{style="color:DarkOrange"} 4.5.3 Cartographic Furniture

Beside map style, **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

In the code chunk below, `tm_compass()`, `tm_scale_bar()` and `tm_grid()` are used to add compass, scale bar and grid lines onto the choropleth map.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

To reset to the default style, the code chunk below can be used:

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tmap_style("white")
```

## [8-1:]{style="color:DarkOrange"} 4.6 Drawing Small Multiple Choropleth Maps

**Small multiple maps**, also referred to as **facet maps**, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In **tmap**, small multiple maps can be plotted in three ways:

-   By assigning multiple values to at least one of the asthetic arguments
-   By defining a group-by variable in *tm_facets()*
-   By creating multiple stand-alone maps with *tmap_arrange()*

### [8-1:]{style="color:DarkOrange"} 4.6.1 Assigning multiple values to at least one of the aesthetic arguments

Small multiple choropleth maps are created by defining ***ncols*** in **tm_fill()**.

::: panel-tabset
## Original

```{r}
#| code-fold: TRUE
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

## Impoved

```{r}
#| code-fold: TRUE
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "brewer.blues") +
  tm_layout(legend.position = c("right", "bottom"),
            legend.width = 7,
            legend.height = 6,
            legend.text.size = 0.65) + 
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```
:::

Small multiple choropleth maps can also be created by assigning multiple values to at least one of the aesthetic arguments.

::: panel-tabset
## Original

```{r}
#| code-fold: TRUE
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

## Improved

```{r}
#| code-fold: TRUE
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"),
            legend.width = 7,
            legend.height = 6,
            legend.text.size = 0.65,
            legend.title.size = 0.7)
```
:::

### [8-1:]{style="color:DarkOrange"} 4.6.2 Defining a group-by variable in *tm_facets()*

Multiple small choropleth maps can be created using **tm_facets()**.

```{r}
#| code-fold: TRUE
#| code-summary: "Show Code"
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

### [8-1:]{style="color:DarkOrange"} 4.6.3 Creating Multiple Stand-Alone Maps with *tmap_arrange()*

In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with `tmap_arrange()`.

::: panel-tabset
## Original

```{r}
#| code-fold: TRUE
#| code-summary: "Show the code"
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## Improved

```{r}
#| code-fold: TRUE
#| code-summary: "Show the code"
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues") +
    tm_layout(legend.position = c("right", "bottom")) 

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues") +
    tm_layout(legend.position = c("right", "bottom")) 

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```
:::

## [8-1:]{style="color:DarkOrange"} 4.7 Mappping Spatial Object Meeting a Selection Criterion

Instead of creating small multiple choropleth map, the selection funtion can be used to map spatial objects meeting the selection criterion.

:::: panel-tabset
## Original

```{r}
#| code-fold: TRUE
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

## Improved

```{r}
#| code-fold: TRUE
#| code-summary: "Show the code"
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE) +
  tm_layout(legend.outside = TRUE,
            legend.outside.position = "right",
            legend.width = 0.3,
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

::: callout-note
I tried to fix the histogram scaling but I am unable to do so
:::
::::

# [8-1:]{style="color:DarkOrange"} 5 Reference

## [8-1:]{style="color:DarkOrange"} 5.1 All about tmap Package

-   [tmap: Thematic Maps in R](https://www.jstatsoft.org/article/view/v084i06)
-   [tmap](https://cran.r-project.org/web/packages/tmap/index.html)
-   [tmap: get started!](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)
-   [tmap: changes in version 2.0](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-changes-v2.html)
-   [tmap: creating thematic maps in a flexible way (useR!2015)](http://von-tijn.nl/tijn/research/presentations/tmap_user2015.pdf)
-   [Exploring and presenting maps with tmap (useR!2017)](http://von-tijn.nl/tijn/research/presentations/tmap_user2017.pdf)

## [8-1:]{style="color:DarkOrange"} 5.2 Geospatial Data Wrangling

-   [sf: Simple Features for R](https://cran.r-project.org/web/packages/sf/index.html)
-   [Simple Features for R: StandardizedSupport for Spatial Vector Data](https://journal.r-project.org/archive/2018/RJ-2018-009/RJ-2018-009.pdf)
-   [Reading, Writing and Converting Simple Features](https://cran.r-project.org/web/packages/sf/vignettes/sf2.html)

## [8-1:]{style="color:DarkOrange"} 5.3 Data Wrangling

-   [dplyr](https://dplyr.tidyverse.org/)
-   [Tidy data](https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html)
-   [tidyr: Easily Tidy Data with ‘spread()’ and ‘gather()’ Functions](https://cran.r-project.org/web/packages/tidyr/tidyr.pdf)
