---
title: "[Hands-On Exercise 9-3:]{style='color:DarkOrange'} Heatmap for Visualising and Analysing Multivariate Data"
date: 2025-06-18
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# [9-3:]{style="color:DarkOrange"} 1 Overview

Heatmaps visualise data through variations in colouring. When applied to a tabular format, heatmaps are useful for cross-examining multivariate data, through placing variables in the columns and observation (or records) in rows and colouring the cells within the table.

Heatmaps are good for showing variance across multiple variables, revealing any patterns, displaying whether any variables are similar to each other, and for detecting if any correlations exist in-between them.

This exercise will cover using R to plot static and interactive heatmaps for visualising and analysing multivariate data.

# [9-3:]{style="color:DarkOrange"} 2 Install and Launch R Packages

The code below installs and launch **seriation**, **heatmaply**, **dendextend** and **tidyverse** in RStudio.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
pacman::p_load(seriation, dendextend, heatmaply, tidyverse, patchwork)
```

# [9-3:]{style="color:DarkOrange"} 3 Import and Prepare the Data

The data of [World Happines 2018 report](https://worldhappiness.report/ed/2018/) will be used. The original data set is in Microsoft Excel format and has been extracted and saved in a csv file called **WHData-2018.csv**.

## [9-3:]{style="color:DarkOrange"} 3.1 Import Data

The code chunk below uses **read_csv()** of readr to import WHData-2018.csv into R and parsed it into tibble R data frame format called **wh**.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh <- read_csv("data/WHData-2018.csv")
```

## [9-3:]{style="color:DarkOrange"} 3.2 Prepare Data

The code below changes the rows by country name instead of row number.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
row.names(wh) <- wh$Country
```

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh
```

::: callout-note
The row number has been added beside the country names.
:::

## [9-3:]{style="color:DarkOrange"} 3.3 Transform Data Frame into a Matrix

The data was loaded into a data frame, but it has to be a data matrix to make a heatmap. The code chunk below will be used to transform *wh* data frame into a data matrix.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

::: callout-note
**wh_matrix** is in R matrix format
:::

# [9-3:]{style="color:DarkOrange"} 4 Static Heatmap

These are some of the many R packages and functions that can be used to draw static heatmaps:

-   [heatmap()](https://www.rdocumentation.org/packages/stats/versions/3.6.0/topics/heatmap) of R stats package. It draws a simple heatmap.
-   [heatmap.2()](https://www.rdocumentation.org/packages/gplots/versions/3.0.1.1/topics/heatmap.2) of **gplots** R package. It draws an enhanced heatmap compared to the R base function.
-   [pheatmap()](https://www.rdocumentation.org/packages/pheatmap/versions/1.0.12/topics/pheatmap) of [**pheatmap**](https://www.rdocumentation.org/packages/pheatmap/versions/1.0.12) R package. **pheatmap** package is also known as Pretty Heatmap. The package provides functions to draws pretty heatmaps and provides more control to change the appearance of heatmaps.
-   [**ComplexHeatmap**](https://bioconductor.org/packages/release/bioc/html/ComplexHeatmap.html) package of R/Bioconductor package. The package draws, annotates and arranges complex heatmaps (very useful for genomic data analysis). The full reference guide of the package is available [here](https://jokergoo.github.io/ComplexHeatmap-reference/book/).
-   [**superheat**](https://cran.r-project.org/web/packages/superheat/) package: A Graphical Tool for Exploring Complex Datasets Using Heatmaps. A system for generating extendable and customisable heatmaps for exploring complex datasets, including big data and data with multiple data types. The full reference guide of the package is available [here](https://rlbarter.github.io/superheat/).

This section will cover plotting static heatmaps by using **heatmap()** of *R Stats* package.

## [9-3:]{style="color:DarkOrange"} 4.1 heatmap() of R Stats

A heatmap will be plotted by using *heatmap()* of Base Stats. The code chunk is given below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

wh_heatmap <- heatmap(wh_matrix,
                      Rowv = NA, Colv = NA,
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

::: callout-note
By default, **heatmap()** plots a cluster heatmap. The arguments **Rowv=NA** and **Colv=NA** are used to switch off the option of plotting the row and column dendrograms.
:::

## [9-3:]{style="color:DarkOrange"} 4.2 Cluster heatmap() of R Stats

Leaving the code as default will plot a cluster heatmap.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh_heatmap <- heatmap(wh_matrix,
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

::: callout-note
-   The order of both rows and columns are different compare to the native wh_matrix
    -   This is because heatmap does a reordering using clusterisation: it calculates the distance between each pair of rows and columns and tries to order them by similarity.
-   Moreover, the corresponding dendrogram is also provided.
-   Red cells denotes high values, and white small ones.
-   This heatmap is not really informative because the Happiness Score variable have relatively higher values, what makes that the other variables with small values all look the same. - Thus, this matrix needs to be normalised - This is done using the scale argument - It can be applied to rows or to columns
:::

## [9-3:]{style="color:DarkOrange"} 4.3 Normalised heatmap() of R Stats

The code below normalises the matrix column-wise.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

::: callout-note
-   The values are scaled now
-   The **margins** argument is used to ensure that the entire x-axis labels are displayed completely and, cexRow and cexCol arguments are used to define the font size used for y-axis and x-axis labels respectively
:::

# [9-3:]{style="color:DarkOrange"} 5 Create Interactive Heatmap

[**heatmaply**](http://talgalili.github.io/heatmaply/index.html) is an R package for building interactive cluster heatmap that can be shared online as a stand-alone HTML file. It is designed and maintained by Tal Galili.

Please refer to [Introduction to Heatmaply](https://cran.r-project.org/web/packages/heatmaply/vignettes/heatmaply.html) to have an overall understanding of the features and functions of Heatmaply package. The [user manual](https://cran.r-project.org/web/packages/heatmaply/heatmaply.pdf) of the package is also useful for reference purposes.

This section will cover using **heatmaply** to design an interactive cluster heatmap with wh_matrix as the input data.

## [9-3:]{style="color:DarkOrange"} 5.1 Work with heatmaply

```{r}
#| code-fold: true
#| code-summary: "Show Code"
heatmaply(mtcars)
```

The code chunk below shows the basic syntax needed to create an interactive heatmap by using the **heatmaply** package.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7)
```

::: callout-note
-   Different from *heatmap()*, for *heatmaply()* the default horizontal dendrogram is placed on the left of the heatmap
-   On the other hand, the text label of each raw is placed on the right side of the heat map
-   When the x-axis marker labels are too long, they will be rotated by 135 degree from the north.
:::

## [9-3:]{style="color:DarkOrange"} 5.2 Data Transformation

When analysing multivariate data sets, it is very common for the variables in the data sets to include values that reflect different types of measurements. In general, these variables have their own range of values. In order to ensure that all the variables have comparable values, data transformation is commonly done before clustering.

Three main data transformation methods are supported by *heatmaply()*; - Scale - Normalize - percentize

### [9-3:]{style="color:DarkOrange"} 5.2.1 Scaling Method

-   When all variables come from or are assumed to come from a normal distribution, then scaling (i.e. subtract the mean and divide by the standard deviation) would bring all the values close to the standard normal distribution.
-   In such a case, each value would reflect the distance from the mean in units of standard deviation
-   The *scale* argument in *heatmaply()* supports column and row scaling.

The code chunk below scales the values of the variables columnise.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column",
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

### [9-3:]{style="color:DarkOrange"} 5.2.2 Normalising Method

-   When variables in the data come from possibly different (and non-normal) distributions, the normalize function can be used to bring data to the 0 to 1 scale by subtracting the minimum and dividing by the maximum of all observations.
-   This preserves the shape of each variable’s distribution while making them easily comparable on the same “scale”.

Different from Scaling, the normalise method is performed on the input data set i.e., *wh_matrix* as shown in the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

### [9-3:]{style="color:DarkOrange"} 5.2.3 Percentizing method

-   This is similar to ranking the variables, but instead of using the rank values, they are divided by the maximal rank
-   This is done by using the ecdf of the variables on their own values, bringing each value to its empirical percentile
-   The benefit of the percentize function is that each value has a relatively clear interpretation, it is the percent of observations is that value or below it.

Similar to Normalize method, the Percentize method is also performed on the input data set i.e. wh_matrix as shown in the code chunk below.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]),
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## [9-3:]{style="color:DarkOrange"} 5.3 Clustering Algorithm

**heatmaply** supports a variety of hierarchical clustering algorithm. The main arguments provided are:

-   *distfun*: Used to compute the distance (dissimilarity) between both rows and columns.
    -   Defaults to dist
    -   The options “pearson”, “spearman” and “kendall” can use correlation-based clustering, which uses as.dist(1 - cor(t(x))) as the distance metric (using the specified correlation method)
-   *hclustfun*: Used to compute the hierarchical clustering when *Rowv* or *Colv* are not dendrograms.
    -   Defaults to *hclust*
-   *dist_method*: default is NULL, which results in using “euclidean”
    -   It can accept alternative character strings indicating the method to be passed to distfun.
    -   By default *distfun* is “dist”” hence this can be one of “euclidean”, “maximum”, “manhattan”, “canberra”, “binary” or “minkowski”
-   *hclust_method*: default is NULL, which results in “complete” method to be used.
    -   It can accept alternative character strings indicating the method to be passed to *hclustfun*.
    -   By default hclustfun is hclust hence this can be one of “ward.D”, “ward.D2”, “single”, “complete”, “average” (= UPGMA), “mcquitty” (= WPGMA), “median” (= WPGMC) or “centroid” (= UPGMC).

In general, a clustering model can be calibrated either manually or statistically.

## [9-3:]{style="color:DarkOrange"} 5.4 Manual Approach

In the code chunk below, the heatmap is plotted using hierachical clustering algorithm with “Euclidean distance” and “ward.D” method.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D",
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## [9-3:]{style="color:DarkOrange"} 5.5 Statistical Approach

The *dend_expend()* and *find_k()* functions of **dendextend** package will be used to determine the best clustering method and number of clusters.

First, the *dend_expend()* will be used to determine the recommended clustering method.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

The output table shows that “average” method should be used because it gave the highest optimum value.

Next, *find_k()* is used to determine the optimal number of clusters.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

The figure above shows that K = 3 would give the best results.

The code below is prepared to incorporate the statistical analysis results above.

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## [9-3:]{style="color:DarkOrange"} 5.6 Seriation

One of the problems with hierarchical clustering is that it does not actually place the rows in a definite order, it merely constrains the space of possible orderings.

-   Take three items A, B and C. If you ignore reflections, there are three possible orderings: ABC, ACB, BAC. If clustering them gives you ((A+B)+C) as a tree, you know that C cannot end up between A and B, but it does not tell you which way to flip the A+B cluster. It does not tell you if the ABC ordering will lead to a clearer-looking heatmap than the BAC ordering.

**heatmaply** uses the seriation package to find an optimal ordering of rows and columns.

-   Optimal means to optimize the Hamiltonian path length that is restricted by the dendrogram structure. This, in other words, means to rotate the branches so that the sum of distances between each adjacent leaf (label) will be minimized. This is related to a restricted version of the travelling salesman problem.

Seriation Algorithm:

1.  Optimal Leaf Ordering (OLO)
    -   This algorithm starts with the output of an agglomerative clustering algorithm and produces a unique ordering, one that flips the various branches of the dendrogram around so as to minimise the sum of dissimilarities between adjacent leaves. The first tab shows the result of applying the Optimal Leaf Ordering to the same clustering result as the heatmap above.
    -   The default options is “OLO” (Optimal leaf ordering) which optimizes the above criterion (in O(n\^4)).
2.  Gruvaeus and Wainer (GW)
    -   Aims for the same goal but uses a potentially faster heuristic
3.  Mean
    -   Gives the output we would get by default from heatmap functions in other packages such as gplots::heatmap.2
4.  None
    -   Gives the dendrograms without any rotation that is based on the data matrix.

::: panel-tabset
## OLO

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO",
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## GW

```{r}
#| code-fold: true
#| code-summary: "Show Code"
#| fig-width: 2
#| fig-height: 4
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW",
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## Mean

```{r}
#| fig-width: 2
#| fig-height: 4
#| code-fold: true
#| code-summary: "Show Code"
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean",
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## None

```{r}
#| fig-width: 2
#| fig-height: 4
#| code-fold: true
#| code-summary: "Show Code"
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)

```
:::

## [9-3:]{style="color:DarkOrange"} 5.7 Working with Colour Palettes

The default colour palette uses by **heatmaply** is *viridis*. heatmaply users, can use other colour palettes in order to improve the aesthetics and visuals of the heatmap.

In the code chunk below, the Blues colour palette of rColorBrewer is used

```{r}
#| fig-width: 2
#| fig-height: 4
#| code-fold: true
#| code-summary: "Show Code"
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 7,
          Colv=NA)
```

## [9-3:]{style="color:DarkOrange"} 5.8 The Finishing Touch

Beside providing a wide collection of arguments for meeting the statistical analysis needs, *heatmaply* also provides many plotting features to ensure a cartographic quality heatmap can be produced.

In the code chunk below the following arguments are used:

-   *k_row* is used to produce 5 groups
-   *margins* is used to change the top margin to 60 and row margin to 200
-   *fontsizw_row* and *fontsize_col* are used to change the font size for row and column labels to 4.\
-   *main* is used to write the main title of the plot
-   *xlab* and *ylab* are used to write the x-axis and y-axis labels respectively

```{r}
#| fig-width: 2
#| fig-height: 4
#| code-fold: true
#| code-summary: "Show Code"

heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```